---
title: "Project Gutenberg Classic Books Analysis"
author: "Denitsa Vasileva"
date: '2020-03-02'
output: html_document
---
# Load packages
```{r, warning=FALSE}
library (here)
library (ggplot2)
library (dplyr)
```

# Introduction 

Project Gutenberg is a free online source which provides free access to more than 60,000
books- mostly classics. Its purpose is to crete digital copies of books in the public domain and
thus make them more accessible and foster a love of reading to new generations of readers.
The Gutenberg project stores troves of information about each available book- including both metadata
about the author and work itself as well as popularity, difficulty and readibility metrics for each book.

# Data Set

The data set which will be used in this project can be downloaded [here](https://corgis-edu.github.io/corgis/csv/classics/?fbclid=IwAR0u9jexVHzWWvP65r7sKXYw8VvKiih80BQIzPVu8HjcpsXwjxVN9n5jCeE).
The rows represent books vailable through Project Gutenberg and the columns are attributes
of these books- e.g. meta data and popularity metrics.

# Exploratory Data Analysis

## Loading File

```{r}
books_data<- read.csv ("classics.csv")
```


## How many observations do we have?

```{r}
dim(books_data)
```
 This shows that in the data there are 38 variables avilable for 1006 books. 
 
## What are the names of the columns (i.e. the vriables we can use in our analysis)?

```{r}
colnames(books_data)
```

# What is the relationship between year of publication and reading difficulty as measured by the automatic readibility index?
```{r}
books_data %>%
  ggplot() +aes (bibliography.publication.year, metrics.difficulty.automated.readability.index) +geom_point() +ggtitle("Plot of Publication Year vs Readbility Difficulty")
  
```

ggplot() +aes ()
