---
title: "Project Gutenberg Classic Books Analysis"
author: "Denitsa Vasileva & Lucy Mosquera"
date: '2020-03-02'
output:
  html_document:
   keep_md: true
---

```{r, include=FALSE}
library(here)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
```

# Introduction 

Project Gutenberg is a free online source which provides free access to more than 60,000
books- mostly classics. Its purpose is to crete digital copies of books in the public domain and
thus make them more accessible and foster a love of reading to new generations of readers.
The Gutenberg project stores troves of information about each available book- including both metadata
about the author and work itself as well as popularity, difficulty and readibility metrics for each book.

# Data Set

The data set which will be used in this project is the classics.csv file which can be downloaded [here](https://corgis-edu.github.io/corgis/csv/classics/?fbclid=IwAR0u9jexVHzWWvP65r7sKXYw8VvKiih80BQIzPVu8HjcpsXwjxVN9n5jCeE).
The rows represent individual books available through Project Gutenberg and the columns are attributes
of these books. The complete set of attributes collected for each book are defined as:

```{r, echo = FALSE}
books_desc <- read.csv(here("Data", "classics_data_description.csv"))

books_desc <- setNames(books_desc, c("Variable", "Data Type", "Description", "Example"))
knitr::kable(books_desc) %>% kable_styling(bootstrap_options = "condensed", full_width = F, fixed_thead = TRUE)
```


# Exploratory Data Analysis

## Loading Data File

```{r}
books_data <- read.csv(here("Data", "classics.csv"))
```


## How many observations do we have?

```{r}
dim(books_data)
```
 This shows that in the data there are 38 variables available for 1006 books. 
 

## What are the names of the columns (i.e. the variables we can use in our analysis)?

```{r}
colnames(books_data)
```


## What is the relationship between year of publication and reading difficulty as measured by the automatic readability index?
```{r}
books_data %>%
  ggplot(aes(bibliography.publication.year, metrics.difficulty.automated.readability.index)) +
  geom_point() + 
  ggtitle("Plot of Publication Year vs Readability Difficulty")
  
```

The graph does not show any significant difference in the level of difficulty in books based on 
Publication Year.


## What is the relationship between average sentence length over time? Has the average gotten longer or shorter over time?

```{r}
books_data %>%
  ggplot(aes(bibliography.publication.year, metrics.statistics.average.sentence.length)) +
  geom_point() +
  facet_wrap(vars())
  ggtitle("Plot of Publication Year vs Sentence Length")
```

The graph does not show any significant difference in the average length of sentences in books based on 
Publication Year.

## Sentiment Analysis

Provided in the Project Gutenberg data is a sentiment polarity score that aims to quantify the positivity or negativity of a work in general. 

```{r}
books_data %>%
  ggplot(aes(bibliography.publication.year, metrics.sentiments.polarity)) +
  geom_point() + 
  ggtitle("Plot of Publication Year vs Sentiment Polarity")
```

This plot does not show a relationship between publication year and sentiment polarity. 

# Research Question

1. Explore the changes in the prevalent sentiments and subjects in books in Project Gutenberg change based on publication year?
2. Explore the changes in the prevalent sentiments and subjects in books in Project Gutenberg change based on author gender, location,etc.?


# Future Analysis

To answer the questions, we will explore different correlations between the columns with semantic metrics
based on different meta data variables for both the authors and the books. We can use linear models for this
as well as visualization tools such as ggplot graphs available in R.

